### 2015-07-06 narrative/data-description.tex
Adjusting size of figure 1.1
Tweaked figure caption
Adjusted size of figure 1.1

\begin{wrapfigure}{r}{0.7\textwidth}
%\vspace{0ex}
        \begin{centering} 
	\includegraphics[scale=0.77]{graphics/SSIO-bucket.pdf}
        %\vspace{1ex}
        \caption{Data is given as a set of chunks which includes utility which contains the lifetime in a storage layer, and the priority of the layer.}
        \label{fig:ssio-bucket}
        \end{centering}
%        \vspace{1ex}
\end{wrapfigure}

We will design new techniques that users can use as part of a data mover, or
the system can automatically use, to raise or lower these sub-chunks up and
down the storage stack. The SSIO layer will then manage the system metadata so
that the user can transparently access any of these sub-chunks without
knowledge of the layer. This requires that when reading the data, the APIs will
then get a time estimate of reading, and then allow the users to determine if
they will read less (only the highest priority sub-chunks) if the time estimate
for reading exceeds their expectations. Existing Sirocco functionality may be
sufficient for these purposes, but it must still be proven in practice.
%

Users will also need to describe information about the data in order to choose
the ``most optimal'' data-refactoring routine. Currently, we understand that
users can tell us that the data represents: 1) phase-space (e.g., the particles
in our example), 2) spatial-changes (e.g., the mesh), 3) space-time (e.g.,
the fields on the mesh). We need this information to get the best possible data
refactoring methods. Although users can request a user-defined data-refactoring
method, system-level methods can use this information to make better decisions.
We will investigate new techniques and new semantics to take this information
and use this to best prioritize and re-organize information.

Finally we will investigate additional information on top of the data model.
This information will be used to indicate data associations that exceed
traditional data models. For example, the fields and mesh in our example have
an inherent relationship. A common data model will describe all of the
variables in the mesh (the coordinates, the units, and the types of elements if
this were a finite-element mesh). We will explore new mechanisms that will
group data together making it explicit that M1, F1, and G1 need to be organized
together.  In otherwords, in will not be helpful if data from M1 and M2 were
organized together along with only F1. When the users wants to visualize data
from the field, F1, they will need only M1 and M2 would be additional
information. Likewise, if the user wants to see F1 and F2, then they will need
the mesh from M1 and M2. This requires that the time to access all of this
data is approximately the same requiring similar access times for all of the
data.  We will investigate what choices the SSIO system must make in order to
keep these pieces ``together'', from a performance perspective. Can the storage
system move everything to a lower layer when all of it doesn't fit or is there
other, less tightly related data can be evicted to preserve the collective
retrieval performance?  We need to investigate the potential problems when this
occurs and how the users can specify additional information to ensure they get
the best possible data from the SSIO layer.

We seek to provide dynamic runtime information to the user and make the
system more transparent. The end user could analyze or visualize data in a more
predictable way since they have confidence how much longer they have to wait
for the data.

@@

\paragraph{Challenges:}
These new techniques offer both technical and adoption challenges. We will only
consider the technical challenges. The proposed techniques and features aim to
expose broad system level environmental information to the application and
allow the application to adapt dynamically based on this information.  One
aspect of this challenge is the design of an interrogative API that offers a
negotiation between the application and the storage system to make adaptation
decisions.  Since we will work closely with many of the leading LCF
applications, we need to make sure that our additional APIs and semantics in
the storage layer will be accepted by these applications and eventually by the
rest of the community. This requires careful examination of all possible
design choices to ensure stable APIs and semantics that do not burden the users
with confusing APIS and semantics that cannot even be used by the system.

>>>>

\begin{wrapfigure}{r}{0.7\textwidth}
%\vspace{0ex}
        \begin{centering} 
	\includegraphics[scale=0.75]{graphics/SSIO-bucket.pdf}
        %\vspace{1ex}
        \caption{The system manages the data across the various stages of the data lifecycle}
        \label{fig:ssio-bucket}
        \end{centering}
%        \vspace{1ex}
\end{wrapfigure}

We will design new techniques, available to both users and the storage system,
to raise or lower these sub-chunks up and
down the storage stack. The SSIO layer will then manage the system metadata so
that the user can transparently access any of these sub-chunks without
the need to know where the data currently resides.
This requires that when reading the data, the middleware will
provide a time estimate of reading, and then allow the users to decide if
they will read less (e.g., only the highest priority sub-chunks)
if the time estimate
for reading exceeds their expectations. Existing Sirocco functionality may be
sufficient for these purposes, but it must still be proven in practice.
%

Users will also need to describe information about the data in order to choose
the optimal data-refactoring routine. Currently, we understand that
users can tell us that the data represents: 1) phase-space (e.g., the particles
in our example), 2) spatial-changes (e.g., the mesh), 3) space-time (e.g.,
the fields on the mesh). We need this information to select
the best possible data
refactoring methods. Although users can insert a user-defined data-refactoring
method, system-level methods can also use this information to make better decisions.
We will investigate new techniques and new semantics capable of using
this information to best prioritize and re-organize data.

Finally we will investigate the use of additional information
on top of the data model.
This information will be used to indicate data associations that exceed
traditional data models. For example, the fields and mesh in our example have
an inherent relationship. A common data model will describe all of the
variables in the mesh (the coordinates, the units, and the types of elements if
this were a finite-element mesh). We will explore new mechanisms that will
group data together making it explicit that M1, F1, and G1 need to be organized
together.  In other words, it will not be helpful if data from M1 and M2 were
organized together along with only F1. When the users wants to visualize data
from the field, F1, they will need only M1 and M2 would be additional
information. Likewise, if the user wants to see F1 and F2, then they will need
the mesh from M1 and M2. This requires that the time to access all of this
data is approximately the same requiring similar access times for all of the
data.  We will investigate what choices the SSIO system must make in order to
keep these pieces ``together'', from a performance perspective. Can the storage
system move everything to a lower layer when all of it doesn't fit or is there
other, less tightly related data can be evicted to preserve the collective
retrieval performance?  We need to investigate the potential problems when this
occurs and how the users can specify additional information to ensure they get
the best possible data from the SSIO layer.

We seek to provide dynamic runtime information to the user and make the
system more transparent. This would allow the end user to analyze or
visualize data in a more
predictable way since they would have more confidence of how much time
the data retrieval process would take.

@@

\paragraph{Challenges:}
These new techniques offer both technical and adoption challenges. We will only
consider the technical challenges. The proposed techniques and features aim to
expose broad system level environmental information to the application and
allow the application to adapt dynamically based on this information.  One
aspect of this challenge is the design of an interrogative API that offers a
negotiation between the application and the storage system to make adaptation
decisions.  Since we will work closely with many of the leading LCF
applications, we need to make sure that our additional APIs and semantics in
the storage layer will be accepted by these applications and eventually by the
rest of the community. This requires careful examination of the full range
of possible
design choices to ensure stable APIs and useful semantics.

### 2015-07-06 narrative/evaluation.tex

Part of our evaluation will include how accurate we can set the bounds on
time estimates for applications. We will measure these estimates when data
is on just one level of the storage hierarchy, along with the time estimates
when data chunks are split. We will evaluate how fast data can be discovered
when it is split between the hierarchy, allowing us to better understand how
many pieces data can be split into before it takes too long to access most
of the data. We will also evaluate our techniques to examine how well we can
migrate data and update our server to determine if data might get
out-of-sync between data chunks. We will need to clearly understand how well
this system works on current systems and we will need to investigate this on
the next generation OLCF system (Summitt) which we will have early access,
along with all of the applications we will be investigating.

@@

We will also want to access how well our data migration and eviction
policies work with user-knowledge. Will the overall system be able to
utilize more of the storage hierarchy by understanding these intentions and
creating autonomic rules to self-optimize? Furthermore we will want to see
if users can give us more accurate policies which can then be used for
better system resource utilization. If we have more knowledge we would like
to be able to illustrate that this will translate into higher levels of
system utilization.

We will also test to see the burden of users to place more knowledge into
the system compared to the overall effect of more optimal performance they
will get. We realize that initially it will be a large burden for users to
place initial knowledge into the system, but we will evaluate the results as
users increase the knowledge placed into the system with the overall burden.
Our hypothesis from working with many LCF users is that once they get over
the initial burden of placing this knowledge the burden decreases as more
knowledge is placed into the system, and thus, they will be more inclined to
increase the knowledge into the system as the benefits increase.

@@

How will we quantify the quality of data? We will look at errors after
analytics.

How well does the system utilize all of its resources. We feel the more
knowledge means the better the system can utilize all of the resources.

predictability ... are teh errors bounded,

>>>>

Our evaluation will include how accurately we can set the bounds on
time estimates for applications. We will measure these estimates when data
is on just one level of the storage hierarchy, along with the time estimates
when data chunks are split. We will evaluate how fast data can be discovered
when it is split between the hierarchy, allowing us to better understand how
many pieces data can be split into before it takes too long to access most
of the data. We will also evaluate our techniques to examine how well we can
migrate data and update our server to determine if data might get
out-of-sync between data chunks. We will need to clearly understand how well
this system works on current systems and we will need to investigate this on
the next generation OLCF system (Summitt) which we will have early access,
along with all of the applications we will be investigating.

@@

We will also want to assess how well our data migration and eviction
policies work with user-knowledge. Will the overall system be able to
utilize more of the storage hierarchy by understanding these intentions and
creating autonomic rules to self-optimize? Furthermore we will want to see
if users can give us more accurate policies which can then be used for
better system resource utilization. If we have more knowledge we would like
to be able to illustrate that this will translate into higher levels of
system utilization.

One of the risks we face in our design is the unwillingness of users to add
more knowledge to the system. To better understand the burden on the user we
will attempt to quantify the additional burden on the users in relation to
the additional knowledge required for optimization. 
 We realize that initially it will be a large burden for users to
place initial knowledge into the system, but we will evaluate the results as
users increase the knowledge placed into the system with the overall burden.
Our hypothesis from working with many LCF users is that once they get over
the initial burden of placing this knowledge the burden decreases as more
knowledge is placed into the system, and thus, they will be more inclined to
increase the knowledge into the system as the benefits increase.

@@

% How will we quantify the quality of data? We will look at errors after
% analytics.

% How well does the system utilize all of its resources. We feel the more
% knowledge means the better the system can utilize all of the resources.

%predictability ... are the errors bounded,
